Working from home and digital divides: resilience during the pandemic

### 2021-03-22 paper/v2_taylor_francis/response_reviewers.md

---
output:
  pdf_document: default
bibliography: bibliography.bib
---

Dear Prof. Bian,

@@

We chose the time range to reflect the period of first lockdown in the UK, as we
say in the first paragraph of 3.2:
"We are particularly interested in upload speeds and the frequency of speed tests
over the period from March to May $2020$, as government statements indicate this
encompasses the period when UK workers were first told to work from home if at all
possible [@GovUK2020]".

We have also added the following text:
"Schools and various retail, leisure and hospitality businesses were closed from late
March, and restrictions were gradually eased from late May".

> 5. Would suggest a review of the spatiotemporal challenges of the covid-19.

We have now added the following paragraph in the 'Digital divides and economic resilience'
section.

"Currently there is a hotly debated discussion in the literature regarding these
exact changes in transportation and the structure of cities. We now have enough
hard data to observe the drastic change in the space-time geography of cities around
the world [e.g. @google2020; @shibayama2021impact].
Although there is a broad agreement that these changes
during the pandemic have played a pivotal role in stopping the spread of the virus
[@jia2020population; @yang2020taking; @mu2020interplay],
we still have no evidence to what extend these increased levels of working from
home, and the consequent decrease in commuting flows and the structure of cities,
will remain post-pandemic".

@@

> 1. The paper used k-means, a popular unsupervised classification in this study. The optimal number of clusters was determined as 13 based on cluster validation indices (CVIs). While this number could be well justified, I feel it is too high to properly interpret the results. The interpretation and knowledge discovery from clustering results with a high k tend to be tedious. Being affected by the high k, the Results section appears to be very difficult to read and follow. As there are many different CVIs, I feel a lower k will benefit the interpretation of results.

Please see our response to the next point.

> 2. Another major comment is on the multinomial logit regression (MLR). This is partly related to my first comment on a larger k. While there is no hard limit on the number of categories in MLR, the interpretation of results will be difficult with a high k as well. In addition, if the dependence among independent variables (collinearity) occurs, the estimation of the multinomial logit model parameters becomes inaccurate. According to Table 2, it is very likely some of the variables could present collinearity (e.g., distance to nearest metro and tech jobs).

Following your suggestion, we have now opted to present the cluster analysis and
the subsequent multinomial regression using $k = 13$. We indeed believe that the
results are more clear now. we have added the following justification in the Methods
and Data section:

"We initially run the algorithm for $k \in \mathbb{N} \bigcap [5,15]$, calculated
the cluster validity indices (CVIs) and then run the
subsequent multinomial regression -- see the end of this section for more details
for the different *k*.
Following @sardatime, to identify the optimal *k* we used the majority vote for
the following CVIs: Silhouette (max), Score function (max), Calinski-Harabasz (max),
Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).
Nevertheless, we opted against using the optimal $k = 13$ solution as it was too
large to allow for communicable LAD clusters. Instead we opted for a smaller $k = 9$,
which led to a rather similar spatial pattern and, importantly, to a much higher
R-squared in the subsequent explanatory regression ($0.44$ instead of $0.34$)."

Regarding multicolinearity, the highest correlations were observed between $NVQ4+$
and $professional jobs$ ($0.78$) and $NVQ4+$ and earnings ($0.69$). Hence, we decided
to exclude $NVQ4+$ from the regression.

@@

> 3. Although Fig. 1 and 2 exhibit different speed tests patterns in 2019 and 2020, the data mining outcome lacks a baseline (prior to 2020) to compare with. It is very likely that the clusters in 2019 would be different from that in 2020. It will make the results more robust. Of course, I will not push this given the page limit. But some insights may be helpful in the Discussion (however, a Discussion section is missing).

ET: after we finalise k, we can rerun k-means for 2019 and decide - could also show mean speeds on a map for 2019 - this would be part of the response to reviewer 1 q2 and q6.
**Let's discuss this**

> 4. No discussion was included in this manuscript. I expect some insights into various important issues, such as problem causes, any broadband policies and socioeconomic solutions, equitable development, and research uncertainties.

We have tightened up the results section, and begun this discussion in the last
paragraph of the results section and then expanded the conclusion section to include
a discussion of the results.

@@

> 5.  There are three levels of digital divide. According to the last paragraph in Section 2, only the intersection of level-1 and level-3 was included in this research. I am wondering how level-2 may fit in the general research framework. Could education and demographic variables be used to represent level-2 digital divide?

We clarify that the occupations included in the regression are more about the level-2
digital divide, whereas the average earnings and, particularly during the pandemic
the % furloughed, are indications of level-3 digital divides. We did have education
in the previous submission (NVQ4+), but have removed it due to multicollinearity
-- see response for point 2, which give more indication of not only education, but digital
skills.

@@

> 6. A few suggestions on the figures. Fig 1 and 2 have crowded tick labels for x-axis month and weekday. Fig 3 and 4 may be placed side by side with the same y-axis for easy comparison. Actually, I feel “Figure 3 shows that the high standard deviation” may be inaccurate as Fig 4 appears to exhibit higher variability. Fig 5 maps may include the locations of metropolitan or major cities to help interpretation. The warmer-cooler color scheme may match the population size instead of the ordinal cluster numbers. The current color scheme does not help interpretation at all.

x-axis for Figures 1 and 2 have been amended accordingly.

The temporal profiles of the upload clusters are now included in Figure 3.

Figure 5 now labelled Figure 4: the map has now been redesigned using a different
colour palette to better reflect the discrete nature of the underpinning data --
the time-series clusters. Points and labels of the main UK cities were also added.
